---
title: "`r params$report_title`"
author: "`r params$report_author`"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    self_contained: true
params:
  # Path to the freshly dropped CSV to inspect. This is the ONE knob a user sets.
  data_path: !r here::here("1_data", "master_micro_survey.csv")

  # --- Report Text (editable; leave "" to use the smart default) ---
  report_title: "Pre-flight Data Quality Report"
  report_author: "CERP Analytics"
  source_note: "Generated via CERP Analytics — pre-flight QA (0.00)"
---

<!--
================================================================================
 0.00 — PRE-FLIGHT DATA QUALITY REPORT + TEMPLATE RECOMMENDER
--------------------------------------------------------------------------------
 A "zero template" you render against ANY freshly dropped CSV *before* picking a
 visual. It profiles the columns, flags the usual field-data landmines, and
 recommends which production templates (3.01–3.18) the data can actually feed —
 including which column maps to which *_var param.

 This template is PER-DATASET, not per-config: it is deliberately NOT wired into
 the master knit (2.2). Render it on its own, from the project root in RStudio:

     rmarkdown::render(
       here::here("3_templates", "0.00_data_quality_report.Rmd"),
       params     = list(data_path = here::here("1_data", "master_micro_survey.csv")),
       output_dir = here::here("4_output")
     )

 Point `data_path` at any CSV in 1_data/. Field data (anything other than
 master_*.csv) stays local — 4_output/ and 3_templates/*.html are gitignored, so
 a babychecker QA report never reaches git, and nothing here calls any API.
================================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,          # QA report: show findings, not code
  warning = FALSE,
  message = FALSE,
  dev.args = list(bg = "#FAF9F6"),
  fig.width = 9,
  fig.height = 4,
  fig.retina = 3
)
```

```{r load-theme-and-profile}
library(tidyverse)

source(here::here("2_R", "2.0_setup_theme.R"), local = TRUE)  # palette + theme_cerp()
source(here::here("2_R", "2.5_helpers.R"), local = TRUE)       # cerp_profile(), etc.

# DELIBERATE EXCEPTION to the "always cerp_load()" convention: this report exists
# to surface the very problems cerp_load() silently scrubs (header/value
# whitespace), so it must see the RAW file. cerp_profile() only describes — it
# never mutates — so reading raw here is safe.
if (!file.exists(params$data_path)) {
  stop("Data file not found: ", params$data_path,
       "\nCheck the 'data_path' parameter.", call. = FALSE)
}
raw  <- readr::read_csv(params$data_path, show_col_types = FALSE, trim_ws = FALSE)
prof <- cerp_profile(raw)
cols <- prof$columns
ds   <- prof$dataset
```

```{r callout-style, results='asis'}
# House-style callout boxes. Colors are pulled from the live palette (cerp_cols,
# sourced above) so they track theme_colors.yml rather than re-hardcoding hexes.
cat(sprintf('
<style>
  .cerp-callout { padding: 0.7rem 1.05rem; margin: 0.55rem 0; border-radius: 6px;
                  border-left: 4px solid %1$s; background: %2$s; color: %3$s;
                  font-family: "Libre Franklin", sans-serif; font-size: 0.94rem; }
  .cerp-callout.warn { border-left-color: %4$s; background: #fbf3e3; }
  .cerp-callout.flag { border-left-color: %5$s; background: #fbeae4; }
  .cerp-callout.ok   { border-left-color: %6$s; background: #eaf3ef; }
  .cerp-callout .lbl { font-weight: 600; }

  /* Report tables — thin borders + roomy cells so dense tables read cleanly.
     Colors track the live palette (text + box highlight). Inline per-column
     alignment from kable (e.g. right-aligned numerics) overrides the defaults. */
  .cerp-table table { border-collapse: collapse; width: 100%%; margin: 0.6rem 0 0.2rem;
                      font-family: "Libre Franklin", sans-serif; font-size: 1.02rem;
                      color: %3$s; }
  .cerp-table thead th { text-align: left; font-weight: 600; padding: 0.55rem 0.8rem;
                         border-bottom: 2px solid %1$s; white-space: nowrap; }
  .cerp-table tbody td { padding: 0.5rem 0.8rem; vertical-align: top;
                         border-bottom: 1px solid #e2e8ea; }
  .cerp-table tbody tr:nth-child(even) { background: %2$s; }
  /* Recommender-only: center + emphasize the leading Rank column. */
  .cerp-rec-table tbody td:first-child { text-align: center; color: %1$s; font-weight: 600; }
</style>',
  cerp_cols[["primary"]], cerp_cols[["box"]], cerp_cols[["text"]],
  cerp_cols[["sand"]], cerp_cols[["accent"]], cerp_cols[["sage"]]))
```

This is a **pre-flight check** of `` `r basename(params$data_path)` `` — run *before*
choosing a visualization. It describes each column, flags the common field-data
problems, and recommends which CERP templates the data can feed. It changes
nothing: fix anything below in the source data, then re-run.

## Dimensions & duplicates

```{r dimensions, results='asis'}
cat(sprintf(
  "The dataset has **%s rows** and **%s columns**.\n\n",
  format(ds$n_rows, big.mark = ","), ds$n_cols))

# Dataset-level flags as house-style callouts (only shown when they fire).
if (ds$single_row) {
  cat('<div class="cerp-callout flag"><span class="lbl">Single row.</span> ',
      'One data row is not enough for most charts — expect distributions, ',
      'trends, and group comparisons to be meaningless.</div>\n', sep = "")
}
if (ds$duplicate_rows > 0) {
  cat(sprintf('<div class="cerp-callout warn"><span class="lbl">%s duplicate row(s).</span> ',
              ds$duplicate_rows),
      'Fully identical rows can double-count in aggregates — confirm they are real ',
      'repeated observations, not an export/merge artifact.</div>\n', sep = "")
}
if (length(ds$empty_columns) > 0) {
  cat(sprintf('<div class="cerp-callout warn"><span class="lbl">Empty column(s):</span> %s. ',
              paste(ds$empty_columns, collapse = ", ")),
      'These carry no values and are ignored by the recommender.</div>\n', sep = "")
}
if (length(ds$header_whitespace) > 0) {
  cat(sprintf('<div class="cerp-callout warn"><span class="lbl">Header whitespace:</span> %s. ',
              paste(sprintf("'%s'", ds$header_whitespace), collapse = ", ")),
      'cerp_load() squishes these automatically at render time, but tidy the source ',
      'header to be safe.</div>\n', sep = "")
}
if (!ds$single_row && ds$duplicate_rows == 0 &&
    length(ds$empty_columns) == 0 && length(ds$header_whitespace) == 0) {
  cat('<div class="cerp-callout ok"><span class="lbl">No structural issues.</span> ',
      'No duplicate rows, empty columns, or header whitespace.</div>\n', sep = "")
}
```

## Column profile

```{r profile-table, results='asis'}
profile_display <- cols %>%
  transmute(
    Column        = column,
    Type          = type,
    `Missing %`   = missing_pct,
    `Distinct`    = n_distinct,
    `Whitespace`  = whitespace_issues,
    `Date-parse`  = ifelse(is.na(date_parse_rate), "—",
                           scales::percent(date_parse_rate, accuracy = 1)),
    `Outliers`    = ifelse(is.na(outlier_count), "—", as.character(outlier_count))
  )

cat('<div class="cerp-table">')
cat(knitr::kable(profile_display, format = "html", align = "llrrrrr", row.names = FALSE))
cat('</div>')
```

*Type is a deterministic best guess (numeric / categorical / date / id-like /
text / empty). "Whitespace" counts values with stray leading/trailing/internal
spaces; "Date-parse" is the share of non-missing values `cerp_parse_date()` can
read (shown for text/date columns); "Outliers" counts points beyond 1.5×IQR.*

## Missingness

```{r missingness-chart}
miss_df <- cols %>% filter(missing_pct > 0) %>% arrange(missing_pct)

if (nrow(miss_df) == 0) {
  cat("No missing values in any column.")
} else {
  # Inline house-style bar chart (reuses theme_cerp() + palette — the sensible
  # reuse here; there is no template viz_*() for a QA missingness bar).
  miss_df <- miss_df %>% mutate(column = factor(column, levels = column))
  ggplot(miss_df, aes(x = missing_pct, y = column)) +
    geom_col(fill = cerp_cols[["primary"]], width = 0.72) +
    geom_text(aes(label = paste0(missing_pct, "%")),
              hjust = -0.15, family = cerp_font_caption,
              size = 3.2, color = cerp_cols[["subtle"]]) +
    scale_x_continuous(limits = c(0, max(miss_df$missing_pct) * 1.12),
                       expand = expansion(mult = c(0, 0))) +
    labs(title = "Missing values by column",
         subtitle = "Percent of rows with no usable value",
         x = NULL, y = NULL,
         caption = params$source_note) +
    theme_cerp() +
    theme(panel.grid.major.y = element_blank())
}
```

## Per-column issue callouts

```{r column-callouts, results='asis'}
any_issue <- FALSE

for (i in seq_len(nrow(cols))) {
  r <- cols[i, ]
  notes <- character(0)

  if (isTRUE(r$type == "empty")) {
    notes <- c(notes, "column is entirely empty (no usable values)")
  }
  if (!is.na(r$missing_pct) && r$missing_pct >= 20) {
    notes <- c(notes, sprintf("%.1f%% missing — thin coverage for this field",
                              r$missing_pct))
  }
  if (isTRUE(r$whitespace_issues > 0)) {
    notes <- c(notes, sprintf("%d value(s) have stray whitespace (cerp_load scrubs this at render)",
                              r$whitespace_issues))
  }
  if (!is.na(r$date_parse_rate) && r$date_parse_rate > 0 && r$date_parse_rate < 0.9) {
    notes <- c(notes, sprintf("only %s of values parse as dates — mixed/ambiguous date formats",
                              scales::percent(r$date_parse_rate, accuracy = 1)))
  }
  if (!is.na(r$outlier_count) && r$outlier_count > 0) {
    notes <- c(notes, sprintf("%d value(s) beyond 1.5×IQR — inspect for entry errors or true extremes",
                              r$outlier_count))
  }

  if (length(notes) > 0) {
    any_issue <- TRUE
    cls <- if (isTRUE(r$type == "empty") ||
               (!is.na(r$missing_pct) && r$missing_pct >= 50)) "flag" else "warn"
    cat(sprintf('<div class="cerp-callout %s"><span class="lbl">%s</span> (%s): %s.</div>\n',
                cls, r$column, r$type, paste(notes, collapse = "; ")))
  }
}

if (!any_issue) {
  cat('<div class="cerp-callout ok"><span class="lbl">No column-level issues.</span> ',
      'No high missingness, whitespace, date-parse, or outlier flags.</div>\n', sep = "")
}
```

## Recommended visuals

```{r recommender}
# The deterministic template recommender lives in the engine
# (2_R/2.5_helpers.R :: cerp_recommend) so this report and the phase-12 Shiny
# config-builder share ONE rule table — never duplicated. It maps the column-type
# profile (+ light name hints) to the production templates the data can feed, with
# a suggested column -> *_var mapping and a ranking score. Same input always yields
# the same recommendations. Suggestions only: confirm the mapping in the template
# YAML before rendering.
rec_df <- cerp_recommend(prof, raw)
```

```{r recommender-table, results='asis'}
if (nrow(rec_df) == 0) {
  cat('<div class="cerp-callout flag"><span class="lbl">No confident recommendation.</span> ',
      'The column profile did not match any template rule cleanly — check the ',
      'flags above (empty columns, all-text data, single row) and confirm the ',
      'right file was profiled.</div>\n', sep = "")
} else {
  rec_out <- data.frame(
    Rank             = seq_len(nrow(rec_df)),
    Template         = sprintf("%s — %s", rec_df$id, rec_df$template),
    `Why it fits`    = rec_df$why,
    `Suggested mapping` = rec_df$mapping,
    check.names = FALSE, stringsAsFactors = FALSE)
  cat('<div class="cerp-table cerp-rec-table">')
  cat(knitr::kable(rec_out, format = "html", align = "clll", row.names = FALSE))
  cat('</div>')
}
```

*Recommendations are deterministic suggestions from the column profile — always
confirm the column → `*_var` mapping in the chosen template's YAML header before
rendering. Absence of a template here means the required column roles were not
detected, not that the chart is impossible.*

<p style="font-family:'Libre Franklin',sans-serif; font-size:0.8rem; color:#8a9aa3; margin-top:2rem;">
`r params$source_note`
</p>
